---
title: "Week 5 - The 2-Sample T-test: Testing for Mean Differences"
subtitle: "PSY 150: Lab Exercise "
author: "KEY"
date: "`r Sys.Date()`"
output: 
    html_document
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, eval=TRUE)
```

------------------------------------------------------------------------

Outline

0. Complete the "Check-in Survey"
1. What is effect size? (Cohen’s D)
2. Read & clean IMDB movie rating data
3. Create samples for each comparison (Action vs Comedy, Drama vs Comedy, Romance vs Comedy)
4. Explore distributions by group 
5. Visualize mean differences
6. Run two-sample t-test models
7. Visualize the t-distribution with critical values and observed t-statistic
8. Practice rendering Quarto document 
9. Create work space to start your student-led projects (& Assignment-2)

------------------------------------------------------------------------

### Complete Check-in Survey

Answer questions thoughtfully to help me best support you this quarter!

**LINK** (copy this into your web browser): https://forms.gle/fMg39Qrtf3qvbksx8

------------------------------------------------------------------------

### What is an `effect size`?

> An effect size tells us how big a difference is on a standardized scale, so we can compare results across contexts.

Cohen's (1988) effect size recommendations (Cohen's D):

* small  ≈ 0.2
* medium ≈ 0.5
* large  ≈ 0.8

> These cutoffs are rules of thumb- not hard rules. The meaning of an effect **always** depends on research context! 

------------------------------------------------------------------------

### Load packages & plot functions

------------------------------------------------------------------------

```{r}

library(tidyverse)
library(janitor)
library(report)
library(ggplot2movies)
library(gt)

source("load_plot_functions.R")

```

------------------------------------------------------------------------

### Read-in data, clean, & create variables

**Acknowledgement:** The code for this lab exercise was adapted from and inspired by materials authored by Andrew Heiss

> Source reference: (https://github.com/andrewheiss/diff-means-half-dozen-ways?tab=readme-ov-file#data)

------------------------------------------------------------------------

```{r}

movies_clean <- movies %>%
  filter(***) %>% 
  filter(str_detect(title, "^[A-Za-z]")) %>% 
  select(***) 

```


Create 3 samples to compare ratings across each movie type comparison:

1. Drop movies (rows) that were not of the specified genre or were not exclusive
2. Group by genre (`action_comedy`, `drama_comedy`, `romance_comedy`)
3. Randomly sample 200 movies from each group (N_total = 400)

```{r}

set.seed(51234)  # Set seed so that everyone ends up with same samples

ac_movies <- ***

dc_movies <- movies_clean %>% 
  drop_na(drama_comedy) %>%
  group_by(drama_comedy) %>%
  slice_sample(n = 200) %>%
  ungroup()

rc_movies <- movies_clean %>% 
  drop_na(romance_comedy) %>%
  group_by(romance_comedy) %>%
  slice_sample(n = 200) %>%
  ungroup()

```

------------------------------------------------------------------------

### Action v. Comedy: Which genre of movie gets higher ratings?

------------------------------------------------------------------------

Let's take a look at our distributions (grouped by movie type)
```{r}

plot_group_dist(
  data = ac_movies,
  outcome = "rating",
  group_var = "action_comedy",
  title = "Which genre of movie gets higher ratings?",
  subtitle = "Sample of 400 movies from IMDB",
  colors = c("#0288b7", "#a90010"))

```

**Question 1**: Describe the distributions of ratings for action & comedy movies: How are they similar or different? 

**Response:** __________________________________________


Visually check for mean differences by movie type
```{r}

mean_diff_barplot(
  data = ac_movies,
  outcome = "rating",
  group_var = "action_comedy",
  title = "Average Movie Rating",
  colors = c("#0288b7", "#a90010"))

```

**Question 2**: Is the difference in average ratings practically meaningful? Explain why or why not, using the plot and your own judgment:

**Response:** __________________________________________

Estimate a Two Sample T-test model (Action v. Comedy)
```{r}

report(t.test(*** ~ ***, data = ac_movies, var.equal = TRUE))

```

**Question 3**: Explain effect size (Cohen's d) in your own words: 

- Why is it useful? 
- What does it tell us about the difference in ratings for this example?

**Response:** __________________________________________


Plot the estimated *t-statistic* on the *t-distribution* 
```{r}

plot_tdist(t_obs = -4.48, df = 398, alpha = 0.05)

```

**Question 4**: 

- What factor changes the shape of the t-distribution? 
- What percent of the t-distribution lies outside the critical values (dotted lines)?
- Interpret what the meaning of the estimated t-statistic (red line): What does it tell us about the difference between groups?


**Response:** __________________________________________

------------------------------------------------------------------------

### Drama V. Comedy: Which genre of movie gets higher ratings?

------------------------------------------------------------------------

Let's take a look at our distributions (grouped by movie type)
```{r}

plot_group_dist(
  data = dc_movies,
  outcome = "rating",
  group_var = "drama_comedy",
  title = "Which genre of movie gets higher ratings?",
  subtitle = "Sample of 400 movies from IMDB",
  colors = c("***", "#a90010"))

```

Visually check mean rating differences between movie genre groups
```{r}

mean_diff_barplot(
  data = dc_movies,
  outcome = "rating",
  group_var = "drama_comedy",
  title = "Average Movie Rating",
  colors = c("***", "#a90010"))

```

Estimate a Two Sample T-test model  (Drama v. Comedy)
```{r}

report(t.test(*** ~ ***, data = dc_movies, var.equal = TRUE))

```

**Question 5**: 

- Is the estimated *t-statistic* above or below the *critical values* when alpha is .05?

**Response:** __________________________________________

Plot the estimated *t-statistic* on the *t-distribution* 
```{r}
plot_tdist(t_obs = ***, df = ***, alpha = ***)
```

**Question 6**: 

- [If you change `alpha` to .01] What percent of the t-distribution lies outside the critical values (dotted lines)?
- Interpret what the meaning of the estimated t-statistic (red line): What does it tell us about the difference between groups?

**Response:** __________________________________________

------------------------------------------------------------------------

### Romance V. Comedy: Which genre of movie gets higher ratings?

------------------------------------------------------------------------

Let's take a look at our distributions (grouped by movie type)
```{r}

plot_group_dist(
  data = rc_movies,
  outcome = "rating",
  group_var = "romance_comedy",
  title = "Which genre of movie gets higher ratings?",
  subtitle = "Sample of 400 movies from IMDB",
  colors = c("***", "#a90010"))

```

Visually check mean rating differences between movie genre groups
```{r}

mean_diff_barplot(
  data = rc_movies,
  outcome = "rating",
  group_var = "romance_comedy",
  title = "Average Movie Rating",
  colors = c("***", "#a90010"))

```

Estimate a Two Sample T-test model (Romance v. Comedy)
```{r}

report(t.test(*** ~ ***, data = rc_movies, var.equal = TRUE))

```

**Question 7**: Is the estimated *t-statistic* above or below the *critical values* when alpha is .05?

**Response:** __________________________________________


Plot the estimated *t-statistic* on the *t-distribution* 
```{r}

plot_tdist(t_obs = ***, df = ***, alpha = ***)

```

**Question 8**: Interpret what the meaning of the estimated t-statistic (red line): What does it tell us about the difference between groups?

**Response:** __________________________________________

------------------------------------------------------------------------

### Practice rendering the Quarto document - Create an HTML file displaying all plots & code!

> Try clicking 'Render' in the top menu (Note: All code must be completed to successfully render)

------------------------------------------------------------------------

### Create a work space to begin your student-led projects & Assignment-1

> [AND] Practice cleaning an example data set

LINK (copy this into your web browser): 

https://posit.cloud/spaces/695065/content/11118691

------------------------------------------------------------------------




















